An Approach to Off-line Handwriting Recognition through Segmentation Techniques and Deep Neural Networks

### 2017-12-17 Report/paper.tex
algorithm

\subsection{Binarization}
\label{sec:binartext}

We start with binarizing the text image. To achieve this there are a lot of thresholding algorithms to choose from, but the one we use is a combination of Gaussian blur in combination with Otsu's thresholding algorithm. \cite{Otsu79} 
The Gaussian blur will smoothen the noise in the image and smoothen the edges of the text.
Otsu's tresholding algorithm will deliver adequate result for noisy text documents.
The choise of tresholding algorithm depends mostly on the intended purposes of the algorithm.
Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing images with much noise.

@@

\begin{algorithm}[tb]
    \caption{Contour combination algorithm}
    \label{alg:overlap}
\begin{algorithmic}
    \STATE {\bfseries Input:} $contour1$, $contour2$
    \STATE $xGap$ = calculateXaxisDistance $contour1$ $contour2$
    \STATE $yOverlap$ = calculateYoverlap $contour1$ $contour2$
    %TODO ::\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
    \STATE combineContours $contour1$ $contour2$
\end{algorithmic}
The chosen paramters are:
\\
$averageCharacterCount = 8$ 
\\
$requiredYaxisOverlap = 1/2$
\end{algorithm}

@@

The next step in the process is to extract the character images from the found word images.
Possible approaches and their advantages are discussed by \cite{CharSegm} and \cite{CharSegmOld}.
We choose the word oversegmentation approach for character extraction. Word oversegmentation is achieved through skeletonization and using a vertical projection profile which is discussed in consequently \ref{sec:skel} and \ref{sec:vertproj}. 
After oversegmenting the word, we find the actual splitpoints by removing unnecessary splitpoints as explained in \ref{sec:postprocesscharseg}

\subsection{Preprocessing}
We resize the image to a fixed height of 100 pixels. 
Next we binarize the image using a combination of Gaussian blur and Otsu's method. 
The binarization method is identical to \ref{sec:binartext} but included to allow segmentation of words which are not part of a text. Binarization is essential as it is necessary for skeletonization.

\subsection{Skeletonization}
\label{sec:skel}
Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
allows us to create a vertical projection profile. The skeletonization algorithm used is the scipy adaptation of \cite{zsthinning}. % TODO reference scipy

\subsection{Vertical projection histogram}
\label{sec:vertproj}

Using the skeletonized word image, we now create a vertical projection histogram.
This vertical projection histogram takes the sum of every collumn in the image, counting black pixels as a $1$ value, and white pixels as a $0$ value.

@@

\subsection{Oversegmentation correction}
\label{sec:postprocesscharseg}

In this section we provide techniques to verify the correctness of segmentation points.
\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
In this paper we discuss machine learning based approaches precedented by a few rule based techniques.

>>>>

\subsection{Binarization}
We start with binarizing the text image. To achieve this there are a lot of thresholding algorithms to choose from, but the one we use is a combination of Gaussian blur in combination with Otsu's thresholding algorithm. \cite{Otsu79} 
The Gaussian blur will smoothen the noise in the image and smoothen the edges of the text.
Otsu's tresholding algorithm will deliver adequate result for noisy text documents.
The choise of tresholding algorithm depends mostly on the intended purposes of the algorithm.
Additional noise removal techniques and other preprocessing techniques may be required for a better result when processing images with much noise.

@@

\begin{algorithm}[tb]
    \caption{Contour combination algorithm}
   \label{alg:overlap}
\begin{algorithmic}
    \STATE {\bfseries Input:}  $contour1$, $contour2$, $text$
    \STATE $xGap =$ calculateXaxisDistance $contour1$, $contour2$
    \STATE $yOverlap =$ calculateYoverlap $contour1$, $contour2$
    \STATE $avgWidth =$ getAverageWordWith $text$
    \IF{ 
    $xGap < avgWidth $ \\ $* (1 / (averageCharacterCount * 2))$ \\  {\bfseries and} \\ $yOverlap > min(contour1.height, contour2.height)$ \\ $*  requiredYaxisOverlap $ 
    }
    \STATE combineContours $contour1$ $contour2$
    \ENDIF
\end{algorithmic}
The chosen paramters are:
\\
$averageCharacterCount = 8$ 
\\
$requiredYaxisOverlap = 1/2$
\end{algorithm}

@@

The next step in the process is to extract the character images from the found word images.
Again, there is a choice between several different approaches.
In this paper we focus solely on a word oversegmentation approach to character extraction, through skeletonization and using a vertical projection profile. The reasoning begind this can be found in the following papers \cite{CharSegm} and \cite{CharSegmOld}.

After first oversegmenting the word, we will recover the original character through postprocessing steps combining different word segments.

\subsection{Normalization}
The first action is again normalizing the word image.
We will resuze the image to a fixed height of $100px$.
This will result in more consistency in rule-based post processing steps.

\subsection{Binarization}
The word image is binarized.
This is again done with a combination of Gaussian blur and Otsu's method.
The binarization is needed for the next step, because we can only apply a skeletonization algorithm to a binary image.

\subsection{Skeletonization}
Now we will skeletonize our word image.
This will normalize the thickness of the strokes to a single pixel.
This allows us to execute our Verical projection profile and interpret the results easily.
The skeletonization algorithm used is the Zhang Suen thinning algorithm \cite{zsthinning}.

\subsection{Vertical projection histogram}
Using the skeletonized word image, we now create a vertical projection histogram.
This vertical projection histogram takes the sum of every collumn in the image, counting black pixels as a $1$ value, and white pixels as a $0$ value.

@@

\subsection{Oversegmentation correction}
In this section we provide techniques to verify the correctness of segmentation points.
\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
